\section{The \apsq Framework}
\label{sec:framework}
This section details the technical implementation of the \apsq framework and is mostly intended to provide insight into the gearbox to potential developers and interested users.
The framework consists of the following four main components that together form \apsq:
\begin{enumerate}
\item \textbf{Core}: The core contains the internal logic to initiate the modules, to provide the geometry, to facilitate module communication and to run the event sequence.
The core keeps its dependencies to a minimum (it only relies on ROOT) and remains independent from the other components as far as possible.
It is the main component discussed in this section.
\item \textbf{Modules}: A module is a set of methods which execute a part of the simulation chain.
Modules are build as separate libraries and loaded dynamically on demand by the core.
The available modules and their parameters are discussed in detail in Section~\ref{sec:modules}.
\item \textbf{Objects}: Objects form the data passed around between modules using the message framework provided by the core.
Modules can listen and bind to messages with objects they wish to receive.
Messages are identified by the object type they are carrying, but they can also be named to allow redirecting data to specific modules facilitating more sophisticated simulation setups.
Messages are meant to be read-only and a copy of the data should be made if a module wishes to change the data.
All objects are compiled into a separate library which is automatically linked to every module.
More information about the messaging system and the supported objects can be found in Section~\ref{sec:objects_messages}.
\item \textbf{Tools}: \apsq provides a set of header-only 'tools' providing access to common logic shared by various modules.
Examples are the Runge-Kutta solver implemented using the Eigen3 library and the set of template specializations for ROOT and Geant4 configurations.
More information about the tools can be found in Section~\ref{sec:additional_tools_resources}.
This set of tools is different from the set of core utilities the framework provides itself, which is part of the core and explained in Section~\ref{sec:logging_utilities}
\end{enumerate}
Finally, \apsq provides an executable which instantiates the core of the framework, receives and distributes the configuration object and runs the simulation chain.

This section is structured as follows.
Section~\ref{sec:arch} provides an overview of the architectural design of the core and describes its interaction with the rest of the \apsq framework.
The different subcomponents such as configuration, modules and messages are discussed in Sections~\ref{sec:config_parameters} to ~\ref{sec:objects_messages}.
Finally, the section closes with a description of the available framework tools in Section~\ref{sec:logging_utilities}.
Some C++ code will be provided in the text, but readers not interested may skip the technical details.

\subsection{Architecture of the Core}
\label{sec:arch}
The core is constructed as a light-weight framework which provides various subsystems to the modules.
It also contains the part responsible for instantiating and running the modules from the supplied configuration file.
The core is structured around five subsystems of which four are centered around a manager and the fifth contain a set of simple general utilities.
The systems provided are:
\begin{enumerate}
\item \textbf{Configuration}: The configuration subsystem provides a configuration object from which data can be retrieved or stored, together with a TOML-like~\cite{tomlgit} parser to read configuration files.
It also contains the \apsq configuration manager which provides access to the main configuration file and its sections.
It is used by the module manager system to find the required instantiations and access the global configuration.
More information is given in Section~\ref{sec:config_parameters}.
\item \textbf{Module}: The module subsystem contains the base class of all \apsq modules as well as the manager responsible for loading and executing the modules (using the configuration system).
This component is discussed in more detail in Section~\ref{sec:module_manager}.
\item \textbf{Geometry}: The geometry subsystem supplies helpers for the simulation geometry.
The manager instantiates all detectors from the detector configuration file.
A detector object contains the position and orientation linked to an instantiation of a particular detector model.
The detector model contains all parameters describing the geometry of the detector.
More details about geometry and detector models is provided in Section~\ref{sec:models_geometry}.
\item \textbf{Messenger}: The messenger is responsible for sending objects from one module to another.
The messenger object is passed to every module and can be used to bind to messages to listen for.
Messages with objects are also dispatched through the messenger as described in Section~\ref{sec:objects_messages}.
\item \textbf{Utilities}: The framework provides a set of utilities for logging, file and directory access, and unit conversion.
An explanation on how to use of these utilities can be found in Section~\ref{sec:logging_utilities}.
A set of C++ exceptions is also provided in the utilities, which are inherited and extended by the other components.
Proper use of exceptions, together with logging informational messages and reporting errors, make the framework easier to use and debug.
A few notes about the use and structure of exceptions are provided in Section~\ref{sec:error_reporting_exceptions}.
\end{enumerate}

\subsection{Configuration and Parameters}
\label{sec:config_parameters}
Individual modules as well as the framework itself are configured through configuration files.
Explanations on how to use the various configuration files together with several examples have been provided in Section~\ref{sec:configuration_files}.
All configuration files follow the same format, but the way their input is interpreted differs per configuration file.

\subsubsection{File format}
\label{sec:config_file_format}
Throughout the framework, a simplified version of TOML~\cite{tomlgit} is used as standard format for configuration files.
The format is defined as follows:
\begin{enumerate}
\item All whitespace at the beginning or end of a line are stripped by the parser.
In the rest of this format specification the \textit{line} refers to the line with this whitespace stripped.
\item Empty lines are ignored.
\item Every non-empty line should start with either \texttt{\#}, \texttt{[} or an alphanumeric character.
Every other character should lead to an immediate parse error.
\item If the line starts with a hash character (\texttt{\#}), it is interpreted as comment and all other content on the same line is ignored.
\item If the line starts with an open square bracket (\texttt{[}), it indicates a section header (also known as configuration header).
The line should contain a string with alphanumeric characters and underscores, indicating the header name, followed by a closing square bracket (\texttt{]}), to end the header.
After any number of ignored whitespace characters there could be a \texttt{\#} character.
If this is the case, the rest of the line is handled as specified in point~3.
Otherwise there should not be any other character (except the whitespace) on the line.
Any line that does not comply to these specifications should lead to an immediate parse error.
Different section header are allowed to have the same name (unless specified otherwise in the particular configuration description).
All key-value pairs following this section header are part of this section until a new section header is started.
\item If the line starts with an alphanumeric character, the line should indicate a key-value pair.
The beginning of the line should contain a string of alphabetic characters, numbers, dots (\texttt{.}), colons (\texttt{\:}) and underscores (\texttt{\_}), but it may only start with an alphanumeric character.
This string indicates the 'key'.
After a optional number of ignored whitespace, the key should be followed by an equality sign (\texttt{$=$}).
Any text between the \texttt{$=$} and the first \texttt{\#} character not enclosed within a pair of single or double quotes (\texttt{'} or \texttt{"}) is known as the non-stripped string.
Any character after the \texttt{\#} is handled as specified in point 3.
If the line does not contain any non-enclosed \texttt{\#} character, the value ends at the end of the line instead.
The 'value' of the key-value pair is the non-stripped string with all whitespace in front and at the end stripped.
The value may not be empty.
Any line that does not comply to these specifications should lead to an immediate parse error.
\item The value may consist of multiple nested dimensions which are grouped by pairs of square brackets (\texttt{[} and \texttt{]}).
The number of square brackets should be properly balanced, otherwise an error is raised.
Square brackets which should not be used for grouping should be enclosed in quotation marks.
Every dimension is split at every whitespace sequence and comma character (\texttt{,}) not enclosed in quotation marks.
Implicit square brackets are added to the begin and end of the value, if these are not explicitly added.
\item The sections of the value which are interpreted as separate entities are named elements.
For a single value the element is on the zeroth dimension, for arrays on the first dimension and for matrices on the second dimension.
Elements can be forced by using quotation marks, either single or double quotes (\texttt{'} or \texttt{"}).
The number of both types of quotation marks should be properly balanced, otherwise an error is raised.
The conversion to the elements to the actual type is performed when accessing the value.
\item All key-value pairs defined before the first section header are part of a zero-length empty section header.
\end{enumerate}

\subsubsection{Accessing parameters}
\label{sec:accessing_parameters}
Values are accessed via the configuration object.
In the following example, the key is a string called \parameter{key}, the object is named \parameter{config} and the type \parameter{TYPE} is a valid C++ type the value should represent.
The values can be accessed via the following methods:
\begin{minted}[frame=single,framesep=3pt,breaklines=true,tabsize=2,linenos]{c++}
// Returns true if the key exists and false otherwise
config.has("key")
// Returns the number of keys found from the provided initializer list:
config.count({"key1", "key2", "key3"});
// Returns the value in the given type, throws an exception if not existing or a conversion to TYPE is not possible
config.get<TYPE>("key")
// Returns the value in the given type or the provided default value if it does not exist
config.get<TYPE>("key", default_value)
// Returns an array of elements of the given type
config.getArray<TYPE>("key")
// Returns a matrix: an array of arrays of elements of the given type
config.getMatrix<TYPE>("key")
// Returns an absolute (canonical if it should exist) path to a file
config.getPath("key", true /* check if path exists */)
// Return an array of absolute paths
config.getPathArray("key", false /* do not check if paths exists */)
// Returns the value as literal text including possible quotation marks
config.getText("key")
// Set the value of key to the default value if the key is not defined
config.setDefault("key", default_value)
// Set the value of the key to the default array if key is not defined
config.setDefaultArray<TYPE>("key", vector_of_default_values)
// Create an alias named new_key for the already existing old_key or throws an exception if the old_key does not exist
config.setAlias("new_key", "old_key")
\end{minted}

Conversions to the requested type are using the \parameter{from_string} and \parameter{to_string} methods provided by the string utility library described in Section~\ref{sec:string_utilities}.
These conversions largely follow the standard C++ parsing, with one important exception.
If (and only if) the value is retrieved as any C/C++ string type and the string is fully enclosed by a pair of \texttt{"} characters, they are stripped before returning the value.
Strings can thus also be provided with or without quotation marks.

\begin{warning}
    It should be noted that a conversion from string to the requested type is a comparatively heavy operation.
    For performance-critical sections of the code, one should consider fetching the configuration value once and caching it in a local variable.
\end{warning}

\subsection{Modules and the Module Manager}
\label{sec:module_manager}
\apsq is a modular framework and one of its core ideas is to partition functionality in independent modules.
The modules are defined in the subdirectory \textit{src/modules/} in the repository.
The name of the directory is the unique name of the module.
The suggested naming scheme is CamelCase, thus an exemplary module name would be \textit{GenericPropagation}.
There are two different kind of modules which can be defined:
\begin{itemize}
\item \textbf{Unique}: Modules for which always only one single instance runs irrespective of the number of detectors.
\item \textbf{Detector}: Modules which are specific to a single detector.
They are replicated for all required detectors.
\end{itemize}
The type of module determines the constructor used, the internal unique name and the supported configuration parameters.
More details about the instantiation logic for the different types of modules can be found in Section~\ref{sec:module_instantiation}.

\subsubsection{Files of a Module}
\label{sec:module_files}
Every module directory should at minimum contain the following documents (with \texttt{ModuleName} replaced by the name of the module):
\begin{itemize}
\item \textbf{CMakeLists.txt}: The build script to load the dependencies and define the source files of the library.
\item \textbf{README.md}: Full documentation of the module.
\item \textbf{\textit{ModuleName}Module.hpp}: The header file of the module (note that another name can be used for this source file, but that is deprecated).
\item \textbf{\textit{ModuleName}Module.cpp}: The implementation file of the module.
\end{itemize}
The files are discussed in more detail below.
By default, all modules are added to the \textit{src/modules/} directory will be build automatically by CMake.
This means that all subdirectories should feature a module with a \textit{CMakeLists.txt} containing instructions on how to build the respective module.

If a module depends on additional packages which not every user might have installed, one can consider adding the following line to the top of the module's \textit{CMakeLists.txt} (see below):
\begin{minted}[frame=single,framesep=3pt,breaklines=true,tabsize=2,linenos]{cmake}
ALLPIX_ENABLE_DEFAULT(OFF)
\end{minted}
Whether or not this is necessary for a given module will be decided on a case-by-case basis.

General guidelines and instructions for implementing new modules are provided in Section~\ref{sec:building_new_module}.

\paragraph{CMakeLists.txt}
Contains the build description of the module with the following components:
\begin{enumerate}
\item On the first line either \parameter{ALLPIX_DETECTOR_MODULE(MODULE_NAME)} or \parameter{ALLPIX_UNIQUE_MODULE(MODULE_NAME)} depending on the type of the module defined.
The internal name of the module is automatically saved in the variable \parameter{${MODULE_NAME}} which should be used as argument to other functions.
Another name can be used by overwriting the variable content, but in the examples below, \parameter{${MODULE_NAME}} is used exclusively.
\item The following lines should contain the logic to load possible dependencies of the module (below is an example to load Geant4).
Only ROOT is automatically included and linked to the module.
\item A line with \texttt{\textbf{ALLPIX\_MODULE\_SOURCES(\$\{MODULE\_NAME\} \textit{sources})}} defines the module source files. Here, \texttt{sources} should be replaced by a list of all source files relevant to this module.
\item Possibly lines to include additional directories and to link libraries for dependencies loaded earlier.
\item A line containing \parameter{ALLPIX_MODULE_INSTALL(${MODULE_NAME})} to set up the required target for the module to be installed to.
\end{enumerate}

A simple CMakeLists.txt for a module named \parameter{Test} which requires Geant4 is provided below as an example.
\vspace{5pt}

\begin{minted}[frame=single,framesep=3pt,breaklines=true,tabsize=2,linenos]{cmake}
# Define module and save name to MODULE_NAME
# Replace by ALLPIX_DETECTOR_MODULE(MODULE_NAME) to define a detector module
ALLPIX_UNIQUE_MODULE(MODULE_NAME)

# Load Geant4
FIND_PACKAGE(Geant4)
IF(NOT Geant4_FOUND)
    MESSAGE(FATAL_ERROR "Could not find Geant4, make sure to source the Geant4 environment\n$ source YOUR_GEANT4_DIR/bin/geant4.sh")
ENDIF()

# Add the sources for this module
ALLPIX_MODULE_SOURCES(${MODULE_NAME}
    TestModule.cpp
)

# Add Geant4 to the include directories
TARGET_INCLUDE_DIRECTORIES(${MODULE_NAME} SYSTEM PRIVATE ${Geant4_INCLUDE_DIRS})

# Link the Geant4 libraries to the module library
TARGET_LINK_LIBRARIES(${MODULE_NAME} ${Geant4_LIBRARIES})

# Provide standard install target
ALLPIX_MODULE_INSTALL(${MODULE_NAME})
\end{minted}

\paragraph{README.md}
The README.md serves as the documentation for the module and should be written in the Markdown format~\cite{markdown}.
It is automatically converted to \LaTeX using Pandoc~\cite{pandoc} and included in this user manual in Section~\ref{sec:modules}.
By documenting the module functionality in Markdown, the information as also viewable with a web browser at the repository in the module sub-folder.

The README.md should follow the structure indicated in the README file of the DummyModule in \textit{src/modules/Dummy}.
The documentation should contain at least the following sections:
\begin{itemize}
\item The H2-size header with the name of the module and at least the following required elements: the \textbf{Maintainer} and the \textbf{Status} of the module.
If the module is working and well-tested, the status of the module should be \textit{Functional}.
By default, new modules are given the status \textbf{Immature}.
The maintainer entry should mention both the full name and email address of the module maintainer between parentheses.
An example for a minimal header is therefore
\begin{verbatim}
## ModuleName
Maintainer: Example Author (<example@example.org>)
Status: Functional
\end{verbatim}
In addition, the \textbf{Input} and \textbf{Output} objects consumed and dispatched by the module should be mentioned.
\item A H4-size section named \textbf{Description}, containing a short description of the module.
\item A H4-size section named \textbf{Parameters} with all available configuration parameters of the module.
The parameters should be briefly explained in an itemized list with the name of the parameter set as inline code block.
\item A H4-size section with the title \textbf{Usage} which should contain at least one simple example of a valid configuration for the module.
\end{itemize}

\paragraph{\texttt{ModuleName}Module.hpp and \texttt{ModuleName}Module.cpp}
All modules should consist of both a header file and a source file.
In the header file, the module is defined together with all its methods.
Brief Doxygen documentation should be added to explain what every method does.
The source file should provide the implementation of every method and also its more detailed Doxygen documentation.
Method shall only be declared in the header and only defined in the source file to keep the interface clean.

\subsubsection{Module structure}
\label{sec:module_structure}
All modules have to inherit from the \texttt{Module} base class which can be found in \textit{src/core/module/Module.hpp}.
The module base class provides two base constructors, a few convenient methods and several methods to override.
Every module should provide a constructor consuming a fixed set of arguments defined by the framework.
This particular constructor is always called during construction by the module instantiation logic.
The arguments for the constructor differs for unique and detector modules.

\begin{warning}
For both module types, it is \textbf{important to forward the configuration object} to the base class \texttt{Module} using the C++ move semantics instead of creating a local copy.
Otherwise, the framework is not able to retrieve the final and potentially updated configuration settings after finalizing the simulation.
This final configuration is stored in the output data by some modules for later reference.
\end{warning}

For unique modules, the constructor for a \texttt{TestModule} should be:
\begin{minted}[frame=single,framesep=3pt,breaklines=true,tabsize=2,linenos]{c++}
TestModule(Configuration config, Messenger* messenger, GeometryManager* geo_manager): Module(std::move(config)) {}
\end{minted}

For detector modules, the first two arguments are the same, but the last argument is a \texttt{std::shared\_ptr} to the linked detector instead.
It should always forward this detector to the base class together with the configuration object.
Thus, the constructor of a detector module is:
\begin{minted}[frame=single,framesep=3pt,breaklines=true,tabsize=2,linenos]{c++}
TestModule(Configuration config, Messenger* messenger, std::shared_ptr<Detector> detector): Module(std::move(config), std::move(detector)) {}
\end{minted}

The pointer to the Messenger can be used to bind variables to either receive or dispatch messages as explained in~\ref{sec:objects_messages}.
The constructor should be used to bind required messages, set configuration defaults and to throw exceptions in case of failures.
Unique modules can access the GeometryManager to fetch all detector descriptions, while detector modules directly receive the object of their respective detector.

In addition to the constructor, every module can override the following methods:
\begin{itemize}
\item \parameter{init()}: Called after loading and constructing all modules and before starting the event loop.
This method can for example be used to initialize histograms.
\item \parameter{run(unsigned int event_number)}: Called for every event in the simulation run with the event number (starting from one).
An exception should be thrown for every serious error, otherwise an warning should be logged.
\item \parameter{finalize()}: Called after processing all events in the run and before destructing the module.
Typically used to save the output data (like histograms).
Any exceptions should be thrown from here instead of the destructor.
\end{itemize}

Within the \parameter{finalize()} method, every module has access to the full final framework configuration via the \parameter{get\_final\_configuration()} method.
This allows to retrieve and e.g. store the configuration actually used for the simulation alongside the data.
It is not possible to call this method outside the \parameter{finalize()} method.

\subsubsection{Module instantiation}
\label{sec:module_instantiation}
The modules are dynamically loaded and instantiated by the Module Manager.
Modules are constructed, initialized, executed and finalized in the linear order they are defined in the configuration file.
Thus the configuration file should follow the order of the real process.
For every non-special section in the main configuration file (see~\ref{sec:config_parameters} for more details), a corresponding library is searched for which contains the module.
Module library are always named following the scheme \textbf{libAllpixModule\texttt{ModuleName}} reflecting the \texttt{ModuleName} configured via CMake.
The module search order is as follows:
\begin{enumerate}
\item Modules already loaded before from an earlier section header
\item All directories in the global configuration parameter \parameter{library_directories} in the provided order if this parameter exists
\item The internal library paths of the executable, that should automatically point to the libraries that are build and installed together with the executable.
These library paths are stored in RPATH on Linux, see the next point for more information.
\item The other standard locations to search for libraries depending on the operating system.
Details about the procedure Linux follows can be found in~\cite{linuxld}.
\end{enumerate}

If the loading of the module library is successful, it is checked if the module is an unique or a detector module.
The instantiation logic determines a unique name and priority, where a lower number indicates a higher priority, for every instantiation.
The name and priority for the instantiation are determined differently for the two types of modules:
\begin{itemize}
\item \textbf{Unique}: Combination of the name of the module and the \parameter{input} and \parameter{output} parameter (both defaulting to an empty string).
The priority is always zero.
\item \textbf{Detector}: Combination of the name of the module, the \parameter{input} and \parameter{output} parameter (both defaulting to an empty string) and the name of detector this module is executed for.
If the name of the detector is specified directly by the \parameter{name} parameter, the priority is \emph{high}.
If the detector is only matched by the \parameter{type} parameter, the priority is \emph{medium}.
If the \parameter{name} and \parameter{type} are both unspecified and the module is instantiated for all detectors, the priority is \emph{low}.
\end{itemize}
The instantiation logic only allows a single instance for every unique name.
If there are multiple instantiations with the same unique name, the instantiation with the highest priority is kept.
If multiple instantiations with the same unique name and the same priority exist, an exception is raised.

\subsubsection{Parallel execution of modules}
\label{sec:multithreading}
The framework has experimental support for running several modules in parallel.
This feature is disabled by default, it has to supported by the module and it has to enabled by the user as described in Section \ref{sec:framework_parameters}.
A significant speed up can be achieved if the simulation contains multiple detectors or simulates the same module using different parameters.

The framework allows to parallelize the execution of the same type of module, if these would otherwise be executed directly after each other in the linear order.
Thus, as long as the name of the module remains the same, while going through the execution order of all \parameter{run()} methods, all instances are added to a work queue.
The instances are then distributed to a set of worker threads as specified in the configuration or determined from system parameters, which will execute the individual modules.
The module manager will wait for all jobs to finish, before it will continue to handle the next type of module.

To enable parallelization for a module, the following line of code has to be added to the constructor of a module:
\begin{minted}[frame=single,framesep=3pt,breaklines=true,tabsize=2,linenos]{c++}
// Enable parallelization of this module if multithreading is enabled
enable_parallelization();
\end{minted}
By adding this, the module promises that it will work correctly if the run-method is executed multiple times in parallel in separate instantiations.
This means in particular that the module will safely handle access to shared (for example static) variables and it will properly bind ROOT histograms to their directory before the \parameter{run()}-method.
Access to constant operations in the GeometryManager, Detector and DetectorModel is always valid between various threads. Also sending and receiving messages is thread-safe.

\subsection{Geometry and Detectors}
\label{sec:models_geometry}
Simulations are frequently performed for a set of different detectors (such as a beam telescope and a device under test).
All these individual detectors together is what \apsq defines as the geometry.
Every detector has a set of properties attached to it:
\begin{itemize}
\item A unique detector \parameter{name} to refer to the detector in the configuration.
\item The \parameter{position} in the world frame.
This is the position of the geometric center of the sensitive device (sensor) given in world coordinates as X, Y and Z (note that any additional components like the chip and possible support layers are ignored when determining the geometric center).
\item An \parameter{orientation_mode} that determines the way the orientation is applied.
It can be either \texttt{xyz} or \texttt{zxz}, where \texttt{xyz} is used if the parameter is not specified.
The \texttt{xyz} option uses extrinsic Euler angles to apply an rotation around the X-axis, followed by an rotation around the original global Y-axis and finally an rotation around the global Z-axis.
The \texttt{zxz} uses the extrinsic Z-X-Z convention for Euler angles instead (this option is also known as the 3-1-3 or the "x-convention" and the most widely used definition of Euler angles overall~\cite{eulerangles}).
\item The \parameter{orientation} to specify the Euler angles in logical order, interpreted using the method above (or with the \texttt{xyz} method if the \parameter{orientation_mode} is not specified).
\item A \parameter{type} of a detector model, for example \emph{hybrid} or \emph{monolithic}.
The model defines the geometry and parameters of the detector.
Multiple detectors can share the same model.
Several ready-to-use models are shipped with the framework.
\item An optional parameter \parameter{alignment_precision_position} to specify the alignment precision along the three global axes as described in Section~\ref{sec:detector_config}.
\item An optional parameter \parameter{alignment_precision_orientation} for the alignment precision in the three rotation angles as described in Section~\ref{sec:detector_config}.
\item An optional \textbf{electric field} in the sensitive device.
An electric field can be added to a detector by a special module as demonstrated in Section~\ref{sec:module_electric_field}.
\end{itemize}
The detector configuration is provided in the detector configuration file as is explained in Section~\ref{sec:detector_config}.

\subsubsection{Changing and accessing the geometry}
The geometry is needed at a very early stage because it determines the number of detector module instantiations as explained in Section~\ref{sec:module_instantiation}.
The procedure of finding and loading the appropriate detector models is explained in more detail in Section~\ref{sec:detector_models}.

The geometry is directly added from the detector configuration file described in Section~\ref{sec:detector_config}.
The geometry manager parses this file on construction, the detector models are loaded and linked later during geometry closing as described above.
It is also possible to add additional models and detectors directly using \parameter{addModel} and \parameter{addDetector} (before the geometry is closed).
Furthermore it is possible to add additional points which should be part of the world geometry using \parameter{addPoint}.
This can for example be used to add the beam source to the world geometry.

The detectors and models can be accessed by name and type through the geometry manager using \parameter{getDetector} and \parameter{getModel}, respectively.
All detectors can be fetched at once using the \parameter{getDetectors} method.
If the module is a detector-specific module its related Detector can be accessed through the \parameter{getDetector} method of the module base class instead (returns a null pointer for unique modules) as follows:
\begin{minted}[frame=single,framesep=3pt,breaklines=true,tabsize=2,linenos]{c++}
void run(unsigned int event_id) {
    // Returns the linked detector
    std::shared_ptr<Detector> detector = this->getDetector();
}
\end{minted}

\subsubsection{Coordinate systems}

Local coordinate systems for each detector and a global frame of reference for the full setup are defined.
The global coordinate system is chosen as a right-handed Cartesian system with the particle beam along the positive z-axis.
Rotations of the individual devices are performed around the geometrical center of their sensor.

Also the local coordinate systems are right-handed Cartesian systems, with the x- and y-axes defining the sensor plane.
The beam is entering along the positive z-axis.
The origin of the coordinate system is the center of the lower left pixel in the grid, i.e.\ the pixel with the indices (0,0).
This simplifies calculations in the local coordinate system as all positions can either be stated in absolute numbers or in fractions of the pixel pitch.

\subsubsection{Detector models}
\label{sec:detector_models}
Different types of detector models are already available and distributed together with the framework.
These models use the configuration format introduced in Section~\ref{sec:config_file_format} and can be found in the \textit{models} directory of the repository.
Every model extends from the \parameter{DetectorModel} base class which defines the minimum required parameters of a detector model in the framework.
The coordinates place the detector in the global coordinate system. The reference point here is the geometric center of the active sensor.
The number of pixels in the sensor in both x- and y-direction define the active pixel matrix.
The pitch of the individual pixels defines its size and also, together with the number of pixels, the total size of the pixel matrix.
Outside the active pixel matrix, the sensor can feature excess material in all directions in the x-y-plane.
A detector of base class type does not feature a separate readout chip, thus only the thickness of an additional, inactive silicon layer can be specified.
Derived models allow for separate readout chips, possibly connected with bump bonds.

Furthermore, multiple layers of support material can be added to the detector description.
It is possible to place support layers at arbitrary positions relative to the sensor, while the default position is behind the readout chip (or inactive silicon layer).
The support material can be chosen from a set of predefined materials, including PCB and Kapton.

This base detector model can be extended to provide a more detailed geometry.
Currently implemented derived models are the \texttt{MonolithicPixelDetectorModel}, which describes a monolithic detector with all electronics directly implemented in the same silicon wafer as the sensor, and the \texttt{HybridPixelDetectorModel}, which in addition to the features described above also includes a separate readout chip with configurable size and bump bonds between the sensor and the readout chip.

\nlparagraph{Detector model parameters}
Models are defined in configuration files which are used to instantiate the actual model classes.
These files for detector models can contain various types of parameters.
Some are required for all models, others are optional or only supported by certain types of models.
For more details about the steps to perform in order to add and use a new detector model, Section~\ref{sec:adding_detector_model} should be consulted.

The set of base parameters supported by every models is provided below.
These parameters should be given at the top of the file before opening any sections.
\begin{itemize}
\item \parameter{type}: A required parameter describing the type of the model.
At the moment either \textbf{monolithic} or \textbf{hybrid}.
This value determines the supported parameters as discussed later.
\item \parameter{number_of_pixels}: The number of pixels in the 2D pixel matrix.
Determines the base size of the sensor together with the \parameter{pixel_size} parameter below.
\item \parameter{pixel_size}: The pitch of a single pixel in the pixel matrix.
Provided as 2D parameter in the x-y-plane.
This parameter is required for all models.
\item \parameter{sensor_thickness}: Thickness of the active area of the detector model containing the individual pixels.
This parameter is required for all models.
\item \texttt{\textbf{sensor\_excess\_\textit{direction}}}: With direction either \parameter{top}, \parameter{bottom}, \parameter{left} or \textit{right}, where the top, bottom, right and left direction are the positive y-axis, the negative y-axis, the positive x-axis and the negative x-axis, respectively.
Specifies the extra material added to the sensor outside the active pixel matrix in the given direction.
\item \parameter{sensor_excess}: Fallback for the excess width of the sensor in all four directions (top, bottom, left and right).
Used if the specialized parameters described below are not given.
Defaults to zero, thus having a sensor size equal to the number of pixels times the pixel pitch.
\item \parameter{chip_thickness}: Thickness of the readout chip, placed next to the sensor.
\end{itemize}

In addition, multiple layers of support can be added to the detector model.
Every support layer should be defined in its own section headed with the name \texttt{[support]}.
By default, no support layers are added.
Support layers allow for the following parameters.
\begin{itemize}
\item \parameter{size}: Size of the support in 2D (the thickness is given separately below).
This parameter is required for all support layers.
\item \parameter{thickness}: Thickness of the support layers.
This parameter is required for all support layers.
\item \parameter{location}: Location of the support layer.
Either \textit{sensor} to attach it to the sensor (opposite to the readout chip/inactive silicon layer), \textit{chip} to add the support layer behind the chip/inactive layer or \textit{absolute} to specify the offset in the z-direction manually.
Defaults to \textit{chip} if not specified.
If the parameter is equal to \textit{sensor} or \textit{chip}, the support layers are stacked in the respective direction when multiple layers of support are specified.
\item \parameter{offset}: If the parameter \parameter{location} is equal to \textit{sensor} or \textit{chip}, an optional 2D offset can be specified using this parameter, the offset in the z-direction is then automatically determined.
These support layers are by default centered around the middle of the pixel matrix (the rotation center of the model).
If the \parameter{location} is set to \textit{absolute}, the offset is a required parameter and should be provided as a 3D vector with respect to the center of the model (thus the center of the active sensor).
Care should be taken to ensure that these support layers and the rest of the model do not overlap.
\item \parameter{hole_size}: Adds an optional cut-out hole to the support with the 2D size provided.
The hole always cuts through the full support thickness.
No hole will be added if this parameter is not present.
\item \parameter{hole_offset}: If present, the hole is by default placed at the center of the support layer.
A 2D offset with respect to its default position can be specified using this parameter.
\item \parameter{material}: Material of the support. \apsq does not provide a set of materials to choose from, but it is up to the modules using the parameter to implement the materials such that they can use it.
Section~\ref{sec:modules} provides details about the materials supported by the geometry builder module (\parameter{GeometryBuilderGeant4}).
\todo{This should be standardized...}
\end{itemize}

The base parameters described above are the only set of parameters supported by the \textbf{monolithic} model.

The \textbf{hybrid} model adds bump bonds between the chip and the sensor while automatically making sure the chip and support layers are shifted appropriately.
Furthermore, it allows to specify the chip dimensions independently from the sensor size, as the readout chip is treated as separate entity.
The additional parameters for the \textbf{hybrid} model are the following:
\begin{itemize}
\item \texttt{\textbf{chip\_excess\_\textit{direction}}}: With direction either \textit{top}, \textit{bottom}, \textit{left} or \textit{right}.
The chip excess in the specific direction, similar to the \texttt{\textbf{sensor\_excess\_\textit{direction}}} parameter described above.
\item \parameter{chip_excess}: Fallback for the excess width of the chip, defaults to zero and thus to a chip size equal to the dimensions of the pixel matrix.
See the \parameter{sensor_excess} parameter above.
\item \parameter{bump_height}: Height of the bump bonds (the separation distance between the chip and the sensor)
\item \parameter{bump_sphere_radius}: The individual bump bonds are simulated as union solids of a sphere and a cylinder.
This parameter sets the radius of the sphere to use.
\item \parameter{bump_cylinder_radius}: The radius of the cylinder part of the bump.
The height of the cylinder is determined by the \parameter{bump_height} parameter.
\item \parameter{bump_offset}: A 2D offset of the grid of bumps.
The individual bumps are by default positioned at the center of all the single pixels in the grid.
\end{itemize}

\nlparagraph{Accessing specific detector models within the framework}
Some modules are written specific to a particular type of detector model.
In order to ensure receiving a specific detector model from the base class, the model should be downcast.
 The downcast returns a null pointer if the class is not of the appropriate type.
An example for fetching an \texttt{HybridPixelDetectorModel} is the following:
\begin{minted}[frame=single,framesep=3pt,breaklines=true,tabsize=2,linenos]{c++}
// "detector" is a pointer to a Detector object
auto model = detector->getModel();
auto hybrid_model = std::dynamic_pointer_cast<HybridPixelDetectorModel>(model);
if(hybrid_model != nullptr) {
    // The model of this Detector is a HybridPixelDetectorModel
}
\end{minted}

\nlparagraph{Specializing detector models}
A detector model contains default values for all parameters.
Some parameters like the sensor thickness can however vary between different detectors of the same model.
To allow for easy adjustment of these parameters, models can be specialized in the detector configuration file introduced in~\ref{sec:detector_config}.
All model parameters, except the type parameter and the support layers, can be changed by adding a parameter with the exact same key and the updated value to the detector configuration.
The framework will then automatically create a copy of this model with the requested change.

\begin{warning}
Before re-implementing models, it should be checked if the desired change can be achieved using the detector model specialization. For most cases this provides a quick and flexible way to adapt detectors to different needs and setups.
\end{warning}

\nlparagraph{Search order for models}
To support different detector models and storage locations, the framework searches different paths for model files in the following order:
\begin{enumerate}
\item If defined, the paths provided in the global \parameter{models\_path} parameter are searched first.
Files are read and parsed directly.
If the path is a directory, all files in the directory are added (not recursing into subdirectories).
\item The location where the models are installed to (refer to the description of the \parameter{MODEL_DIRECTORY} variable in Section~\ref{sec:cmake_config}).
\item The standard data paths on the system as given by the environmental variable \parameter{$XDG_DATA_DIRS} with ``\project'' appended.
The \parameter{$XDG_DATA_DIRS} variable defaults to \textit{/usr/local/share/} (thus effectively \textit{/usr/local/share/\project}) followed by \textit{/usr/share/} (effectively \textit{/usr/share/\project}).
\end{enumerate}

\subsection{Passing Objects using Messages}
\label{sec:objects_messages}
Communication between modules is established through messages.
Messages are templated instantiations of the \parameter{Message} class carrying a vector of objects.
The list of objects available in the \apsq objects library are discussed in Section~\ref{sec:objects}.
The messaging system has a dispatching mechanism to send messages and a receiving part that fetches incoming messages.

The dispatching module can specify an optional name for the messages, but modules should normally not specify this name directly.
If the name is not given (or equal to \texttt{-}) the \parameter{output} parameter of the module is used to determine the name of the message, defaulting to an empty string.
Dispatching the message to their receivers is then performed following these rules:
\begin{enumerate}
\item The receiving module the will \underline{only} receive a message if it has the exact same type as the message dispatched (thus carrying the same objects).
If the receiver is however listening to the \parameter{BaseMessage} type which does not specify the type of objects it is carrying, it will receive all dispatched messages instead.
\item The receiving module will \underline{only} receive messages with the exact name it is listening for.
The module uses the \parameter{input} parameter to determine to which message names the module should listen.
If the \parameter{input} parameter is equal to \texttt{*} the module should listen to all messages.
Every module by default listens to messages with no name specified (thus receiving the messages of dispatching modules without output name specified).
\item If the receiving module is a detector module, it will \underline{only} receive messages bound to that specific detector \underline{or} messages that are not bound to any detector.
\end{enumerate}

An example how to dispatch a message containing an array of \parameter{Object} types bound to a detector named \texttt{dut} is provided below.
As usual, the message is dispatched at the end of the \parameter{run()} function of the module.
\begin{minted}[frame=single,framesep=3pt,breaklines=true,tabsize=2,linenos]{c++}
void run(unsigned int event_id) {
    std::vector<Object> data;
    // ..fill the data vector with objects ...

    // The message is dispatched only for the module's detector, stored in "detector_"
    auto message = std::make_shared<Message<Object>>(data, detector_);

    // Send the message using the Messenger object
    messenger->dispatchMessage(this, message);
}
\end{minted}

\subsubsection{Methods to process messages}
The message system has multiple methods to process received messages.
The first three are the most common methods and the fourth should only be used if necessary.
\begin{enumerate}
\item Bind a \textbf{single message} to a variable.
This should usually be the preferred method as most modules only expect one message to arrive per event, as a module should typically send only one message containing the list of all relevant objects.
The following example binds to a message containing an array of object and is placed in the constructor of a detector-type \parameter{TestModule}:
\begin{minted}[frame=single,framesep=3pt,breaklines=true,tabsize=2,linenos]{c++}
TestModule(Configuration, Messenger* messenger, std::shared_ptr<Detector>) {
    messenger->bindSingle(this,
                          /* Pointer to the message variable */
                          &TestModule::message,
                          /* No special messenger flags */
                          MsgFlags::NONE);
}
std::shared_ptr<Message<Object>> message;
\end{minted}
\item Bind a \textbf{set of messages} to a \texttt{std::vector} variable.
This method should be used it the module can (and expects to) receive the same message multiple times (possibly because it wants to receive the same type of message for all detectors).
An example to bind multiple messages containing an array of objects in the constructor of a unique-type \parameter{TestModule} would be:
\begin{minted}[frame=single,framesep=3pt,breaklines=true,tabsize=2,linenos]{c++}
TestModule(Configuration, Messenger* messenger, GeometryManager* geo_manager) {
    messenger->bindMulti(this,
                          /* Pointer to the message vector */
                          &TestModule::messages,
                          /* No special messenger flags */
                          MsgFlags::NONE);
}
std::vector<std::shared_ptr<Message<Object>>> messages;
\end{minted}
\item Add a dependency to keep messages in memory for retrieving objects from the history.
Several objects have related objects which can be retrieved as explained in Section \ref{sec:objhistory}.
Linked objects can only be retrieved (during the same event) if the respective source objects remain in memory.
Modules should explicitly mark objects that are accessed indirectly through the history of the messages, to ensure these objects are not deleted.
This method can be used as follows in a detector-specific \parameter{TestModule}, to keep \parameter{PixelHitMessage}s in memory without having direct access to them (\parameter{PixelHitMessage} can be replaced here with any other type of message):
\begin{minted}[frame=single,framesep=3pt,breaklines=true,tabsize=2,linenos]{c++}
TestModule(Configuration, Messenger* messenger, std::shared_ptr<Detector>) {
    messenger->addDependency<PixelHit>(this,
                                       /* No special message flags */
                                       MsgFlags::NONE);
}
\end{minted}
\item Listen to a particular message type and execute a \textbf{listener function} as soon as an object is received.
Can be used for more advanced strategies of retrieving messages, but the other method should be preferred when possible.
The listening module should \underline{not} do any heavy work in the listening function as this is supposed to take place in their \texttt{run} method instead.
Using a listener function can lead to unexpected behavior because the function is executed during the run method of the dispatching module.
This means that logging is performed at the level of the dispatching module and that the listener method can be accessed from multiple threads if the dispatching module is parallelized.
Listening to a message containing an array of objects in a detector-specific \parameter{TestModule} could be performed as follows:
\begin{minted}[frame=single,framesep=3pt,breaklines=true,tabsize=2,linenos]{c++}
TestModule(Configuration, Messenger* messenger, std::shared_ptr<Detector>) {
    messenger->registerListener(this,
                                /* Pointer to the listener method */
                                &TestModule::listener,
                                /* No special message flags */
                                MsgFlags::NONE);
}
void listener(std::shared_ptr<Message<Object>> message) {
    // Do something with the received message ...
}
\end{minted}
\end{enumerate}

\subsubsection{Message flags}
Flags can be added to the bind and listening methods which enable a particular behavior of the framework.
\begin{itemize}
\item \parameter{REQUIRED}: Specifies that this message is required to be received.
If the particular type of message is not received before it is time to execute the module's run function, the execution of the method is automatically skipped by the framework for the current event.
This can be used to ignore modules which cannot perform any action without received messages, for example charge carrier propagation without any deposited charge carriers.
\item \parameter{NO_RESET}: Messages are by default automatically reset after the \parameter{run} function has been executed to prevent older messages from previous events to reappear.
This behavior can be disabled by setting this flag (this does not have any effect for listening functions).
Setting this flag for single bound messages (without \parameter{ALLOW_OVERWRITE}, see below) would cause an exception to be raised if the message is overwritten in a later event.
\item \parameter{ALLOW_OVERWRITE}: By default an exception is automatically raised if a single bound message is overwritten (thus receiving it multiple times instead of once).
This flag prevents this behavior.
It can only be used for variables bound to a single message.
\item \parameter{IGNORE_NAME}: If this flag is specified, the name of the dispatched message is not considered.
Thus, the \parameter{input} parameter is ignored and forced to the value \texttt{*}.
\end{itemize}

\subsection{Logging and other Utilities}
\label{sec:logging_utilities}
The \apsq framework provides a set of utilities which improve the usability of the framework and extend the functionality provided by the C++ Standard Template Library (STL).
The former include a flexible and easy-to-use logging system, introduced in Section~\ref{sec:logger} and an easy-to-use framework for units that supports converting arbitrary combinations of units to common base units which can  be used transparently through the framework and which will be discussed in more detail in Section~\ref{sec:unit_system}.
The latter comprise tools which provide functionality the C++14 standard does not contain.
These utilities are used internally in the framework and are only shortly discussed in Section~\ref{sec:filesystem} (file system support) and Section~\ref{sec:string_utilities} (string utilities).

\subsubsection{Logging system}
\label{sec:logger}
The logging system is build to handle input/output the same way as \texttt{std::cin} and \texttt{std::cout} do.
This approach is both very flexible and easy to read.
The system is globally configured, thus only one logger instance exists.
In order to send a message to the logging system at a level of \parameter{LEVEL}, the following can be used:
\begin{minted}[frame=single,framesep=3pt,breaklines=true,tabsize=2,linenos]{c++}
LOG(LEVEL) << "this is an example message with an integer and a double " << 1 << 2.0;
\end{minted}
A newline is added at the end of every log message.
Multi-line log messages can also be used: the logging system will automatically align every new line under the previous message and will leave the header space empty on new lines.

The system also allows for producing a message which is updated on the same line for simple progress-bar-like functionality.
It is enabled using the \parameter{LOG_PROCESS(LEVEL, IDENTIFIER)} command.
Here, the \parameter{IDENTIFIER} is a unique string identifying this output stream in order not to mix different progress reports.

If the output is a terminal screen the logging output will be colored to make it prettier to read.
This will be disabled automatically for all non-terminal outputs.

More details about the logging levels and formats can be found in Section~\ref{sec:logging_verbosity}.

\subsubsection{Unit system}
\label{sec:unit_system}
Correctly handling units and conversions is of paramount importance.
Having a separate C++ type for all different kind of units would however be too cumbersome for a lot of operations.
Therefore, the units are stored in standard C++ floating point types in a default unit which all code in the framework should use for calculations.
In configuration files as well as for logging it is however very useful to provide quantities in different units.

The unit system allows adding, retrieving, converting and displaying units.
It is a global system transparently used throughout the framework.
Examples of using the unit system are given below:
\begin{minted}[frame=single,framesep=3pt,breaklines=true,tabsize=2,linenos]{c++}
// Define the standard length unit and an auxiliary unit
Units::add("mm", 1);
Units::add("m", 1e3);
// Define the standard time unit
Units::add("ns", 1);
// Get the units given in m/ns in the defined framework unit (mm/ns)
Units::get(1, "m/ns");
// Get the framework unit (mm/ns) in m/ns
Units::convert(1, "m/ns");
// Return the unit in the best type (lowest number larger than one) as string.
// The input is in default units 2000mm/ns and the 'best' output is 2m/ns (string)
Units::display(2e3, {"mm/ns", "m/ns"});
\end{minted}

More details about how the unit system is used within \apsq can be found in Section~\ref{sec:config_values}.

\subsubsection{Internal utilities}
\label{sec:filesystem}
The \textbf{filesystem} utilities provide functions to convert relative to absolute canonical paths, to iterate through all files in a directory and to create new directories.
These functions should be replaced by the C++17 file system API~\cite{cppfilesystem} as soon as the framework minimum standard is updated to C++17.

\label{sec:string_utilities}
The STL only provides string conversions for standard types using \texttt{std::stringstream} and \texttt{std::to\_string} which do not allow to parse strings encapsulated in pairs of double quote (\texttt{"}) characters and neither does it allow to integrate different units.
Furthermore it does not provide wide flexibility to add custom conversions for other external types in either way.

The \apsq \texttt{to\_string} and \texttt{from\_string} methods provied by its \textbf{string utilities} do allow for these flexible conversions and are extensively used in the configuration system.
Conversions of numeric types with a unit attached are automatically resolved using the unit system discussed in Section~\ref{sec:unit_system}.
The string utilities also include trim and split strings functions missing in the STL.

Furthermore, the \apsq tool system contains extensions to allow automatic conversions for ROOT and Geant4 types as explained in Section~\ref{sec:root_and_geant4_utilities}.


\subsection{Error Reporting and Exceptions}
\label{sec:error_reporting_exceptions}
\apsq generally follows the principle of throwing exceptions in all cases something is definitely wrong.
Exceptions are also thrown to signal for errors in the user configuration.
It does not attempt to circumvent problems or correct configuration mistakes.
Error return codes should not be used at all.
The asset of this method is that errors cannot easily be ignored and the code is more predictable in general.

For warnings and informational messages, the logging should be used extensively.
This helps both in following the progress of the simulation and in debugging problems.
Care should however be taken to limit the amount of messages in levels higher than \texttt{DEBUG} or \texttt{TRACE}.
More details about the logging levels and their usage can be found in Section~\ref{sec:logging_verbosity}.

The base exceptions in \apsq are available via the utilities.
The most important exception base classes are the following:
\begin{itemize}
\item \parameter{ConfigurationError}: All errors related to incorrect user configuration.
This could indicate a non-existing configuration file, a missing key or an invalid parameter value.
\item \parameter{RuntimeError}: All other errors arising at run-time.
Could be related to incorrect configuration if messages are not correctly passed or non-existing detectors are specified.
Could also be raised if errors arise while loading a library or executing a module.
\item \parameter{LogicError}: Problems related to modules which do not properly follow the specifications, for example if a detector module fails to pass the detector to the constructor.
These methods should never be raised for correctly implemented modules and should therefore not be triggerable by users.
Reporting this type of errors can help developers during their development of new modules.
\end{itemize}

Outside the core framework, exceptions can also be used directly by the modules.
There are a few exceptions which should typically be used by modules to indicate errors:
\begin{itemize}
\item \parameter{InvalidValueError}: Derived from configuration exceptions.
Signals any problem with the value of a configuration parameter not related to parsing or conversion to the required type.
Can for example be used for parameters where the possible valid values are limited, like the set of logging levels, or for paths that do not exist.
An example is shown below:
\begin{minted}[frame=single,framesep=3pt,breaklines=true,tabsize=2,linenos]{c++}
void run(unsigned int event_id) {
    // Fetch a key from the configuration
    std::string value = config.get("key");

    // Check if it is a 'valid' value
    if(value != 'A' && value != "B") {
        // Raise an error if it the value is not valid
        //   provide the configuration object, key and an explanation
        throw InvalidValueError(config, "key", "A and B are the only allowed values");
    }
}
\end{minted}
\item \parameter{InvalidCombinationError}: Derived from configuration exceptions.
Signals any problem with a combination of configuration parameters used.
This could be used if several optional but mutually exclusive parameters are present in a module, and it should be ensured that only one is specified at the time.
The exceptions accepts the list of keys as initializer list.
An example is shown below:
\begin{minted}[frame=single,framesep=3pt,breaklines=true,tabsize=2,linenos]{c++}
void run(unsigned int event_id) {
    // Check if we have mutually exclusive options defined:
    if(config.count({"exclusive_opt_a", "exclusive_opt_b"}) > 1) {
        // Raise an error if the combination of keys is not valid
        //   provide the configuration object, keys and an explanation
        throw InvalidCombinationError(config, {"exclusive_opt_a", "exclusive_opt_b"}, "Options A and B are mutually exclusive, specify only one.");
    }
}
\end{minted}

\item \parameter{ModuleError}: Derived from module exceptions.
Should be used to indicate any runtime error in a module not directly caused by an invalid configuration value, for example that it is not possible to write an output file.
A reason should be given to indicate what the source of problem is.
\todo{The module class should be passed as well, so the module name can be displayed in the error message}
\end{itemize}

\todo{add more info about error reporting style?}
